---
title: Boundaries
format: gfm
---

```{r include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Local authority districts

Local authority districts are provided by https://geoportal.statistics.gov.uk/.
They are not static

```{r lads}
layer = "Local_Authority_Districts_December_2022_Boundaries_UK_BUC"
u = paste0(
  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/",
  layer,
  "/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
)
lads = sf::read_sf(u)
plot(lads$geometry)
```

## Transport authorities

Transport authorities are composed of one or more local authorities.

```{r, eval=FALSE, echo=FALSE}
u = "https://github.com/udsleeds/openinfra/raw/main/data-small/lad_ta_region_lookup_atf3.csv"
lad_lookup = read_csv(u)
ta_lookup = read_csv("C:/Users/RLOVELAC/atip_analysis/lookup.csv")
summary(ta_lookup$name %in% lad_lookup$Region_name)
for(i in seq(length(ta_lookup))) {
  print(summary(ta_lookup[[i]] %in% lad_lookup$Region_name))
}
ta_in_lookup = ta_lookup$name_original %in% lad_lookup$Region_name
ta_lookup[!ta_in_lookup, ] # South Yorkshire missing
lad_lookup = lad_lookup %>% 
  mutate(ATF4_region_name = case_when(
    str_detect(Region_name, "Sheff") ~ "South Yorkshire Combined Authority",
    TRUE ~ Region_name
      ))
summary(ta_lookup$name_original %in% lad_lookup$ATF4_region_name)
ta_lookup = ta_lookup %>% 
  transmute(
    ATF4_region_name = name_original,
    internal_region_name = name,
    authority_name_alternative_1 = Authority,
    authority_name_alternative_2 = Authority_alternative,
    authority_name_alternative_3 = `LA name`
    )
lad_lookup = lad_lookup %>% 
  transmute(
    LAD22NM,
    ATF4_region_name
    )
write_csv(lad_lookup, "lad_lookup.csv")
write_csv(ta_lookup, "ta_lookup.csv")
```

The lookup table in `lad_lookup.csv` provides a lookup between LADs and transport authorities.
The table has the following contents (sample of 9 authorities shown):

```{r}
lad_lookup = read_csv("lad_lookup.csv")
lad_lookup %>% 
  slice(1:9) %>% 
  knitr::kable()
```


The lookup table was used to create a dataset representing transport authorities, as shown below.

```{r transport_authorities}
lads_joined = left_join(lads, lad_lookup)
lads_joined = lads_joined %>% 
  filter(!is.na(ATF4_region_name))
sf::sf_use_s2(FALSE)
transport_authorities = lads_joined %>% 
  group_by(ATF4_region_name) %>% 
  summarise(
    n_lads = n(),
    lad_names = paste(LAD22NM, collapse = ", ")
  )
plot(transport_authorities$geometry)
sf::write_sf(transport_authorities, "transport_authorities.geojson", delete_dsn = TRUE)
```

There are `r nrow(transport_authorities)` transport authorities including London and Isles of Scilly.


This is shown below for West Yorkshire, for example.

